Red Wine Data Exploration by Shivam Bhardwaj
========================================================
What properties contributes in making of good red wine? In this project we try 
to answer this question by exploring the red wine data set.

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.


library(ggplot2)
library(GGally)
library(scales)
library(memisc)
library(lattice)
library(MASS)
library(gridExtra)
library(dplyr)
library(reshape)
library(corrplot)
library(psych)

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
redwine <- read.csv("wineQualityReds.csv")
```

```{r echo=FALSE,  Remove_Outliers}

# Remove the top 1% for four columns that had the most drastic outliers.
redwine <- redwine[redwine$fixed.acidity < 
                     quantile(redwine$fixed.acidity, 0.99),]
redwine <- redwine[redwine$residual.sugar < 
                     quantile(redwine$residual.sugar, 0.99),]
redwine <- redwine[redwine$total.sulfur.dioxide < quantile(redwine$total.sulfur.dioxide, 0.99),]
redwine <- redwine[redwine$free.sulfur.dioxide < quantile(redwine$free.sulfur.dioxide, 0.99),]

# Remove x column as it's just an index

redwine <- within(redwine, rm(X))

summary(redwine)
str(redwine)
```

Most of the quality ratings are either 5 or 6; with 5 being the most frequent. Quality is a categorical discrete variable, but if we were to treat it as continuous, the mean would be 5.63 and the median would be 6. The highest rating was 8, and the lowest was 3. Additionally, total sulfur dioxide and free sulfur dioxide appeared to be discrete variables.This is likely due to rounding issues.
I would also think that citric acid is a subset of fixed acidity and potentially volatile acidity. 

Fixed acidity,residual sugar, total sulfur dioxide, and free sulfur dioxide were all stripped from their top 1% values as they appeared to be large outliers.

# Univariate Plots Section

This red wine data set contains 1,599 obersvations with 11 variables on the chemical properties of the wine.


```{r echo=FALSE, Univariate_selection}
names(redwine)
``` 

###  Quality Distribution

The wine quality grade is a discrete number. It is ranged from 3 to 8. The 
median value is at 6.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
ggplot(aes(x = quality), data = redwine)+
geom_bar(colour="black", fill="#ff9dab", binwidth = 1)+
                scale_x_continuous()+
                xlab("wine grade") + ylab("count") +
                ggtitle("Distribution of Wine Quality")

```

### Distribution of Other Chemical Properties

Transformed the long tailed sulphates data for a more accurate distribution. The log10 produces a relatively normal distribution, and there is nothing 
particularly striking with the transformations, as given that there are only 
1534 observations being analyzed, it's very likely that many sulphate
measurements won't be included in the data set. Variance decreases for log10 sulphates and graph looks more normal so will keep it.

```{r echo=FALSE, warning=FALSE, Other_Properties}
p1<-
  ggplot(aes(x = fixed.acidity), data = redwine)+
        xlab("fixed acidity")+
        geom_bar(colour="black", fill="#ffb6c1",binwidth = 0.5)
p2<-
  ggplot(aes(x = volatile.acidity), data = redwine)+
        xlab("volatile acidity")+
        geom_bar(colour="black", fill="#ffb6c1",binwidth = 0.05)
p3<-
  ggplot(aes(x = citric.acid), data = redwine)+
        xlab("citric acid")+
        geom_bar(colour="black", fill="#ffb6c1", binwidth = 0.05)
p4<-
  ggplot(aes(x = residual.sugar), data = redwine)+
        xlab("residual sugar")+
        geom_bar(colour="black", fill="#ffb6c1", binwidth = 0.4)
p5<-
  ggplot(aes(x = chlorides), data = redwine)+
        geom_bar(colour="black", fill="#ffb6c1", binwidth = 0.025)
p6<-
  ggplot(aes(x = free.sulfur.dioxide), data = redwine)+
        xlab("free sulfur dioxide")+
        geom_bar(colour="black", fill="#ffb6c1",binwidth = 4)
p7<-
  ggplot(aes(x = total.sulfur.dioxide), data = redwine)+
        xlab("total sulfur dioxide")+
        geom_bar(colour="black", fill="#ffb6c1", binwidth = 20)
p8<-
  ggplot(aes(x = density), data = redwine)+
        geom_bar(colour="black", fill="#ffb6c1", binwidth = 0.001)
p9<-
  ggplot(aes(x = pH), data = redwine)+
        geom_bar(colour="black", fill="#ffb6c1", binwidth = 0.05)
p10<-
  ggplot(aes(x = sulphates), data = redwine)+
        geom_bar(colour="black", fill="#ffb6c1", binwidth = 0.1)
p11<-
  ggplot(aes(x = alcohol), data = redwine)+
        geom_bar(colour="black", fill="#ffb6c1",binwidth = 0.25)

grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,ncol=3)
```

It appears that we can actually group wine quality into three distinct 
categories: bad, average, and excellent. Most of the red wines were average, followed by excellent, and then bad. It seems like the red wines overall were 
very average, with a few having excellent tastes. I'm interested in what makes a wine excellent or bad -- not what makes it average.

```{r echo=FALSE, Rating}

# Convert quality into 3 subqualities: bad, average, excellent. 
redwine$rating <- ifelse(redwine$quality < 5, 'bad', 
                    ifelse(redwine$quality < 7, 'average', 
                           ifelse(redwine$quality > 6, 'excellent', 'none')))

redwine$rating <- ordered(redwine$rating, levels = c('bad', 'average', 'excellent'))
summary(redwine$rating)
ggplot(redwine, aes(rating)) + geom_bar(fill = "#ffb6c1") 
```

# Univariate Analysis

Some observation on the distribution of the chemical property can be made:

- **Normal**:  Volatile acidity, Density, PH

- **Positively Skewed**: Fixed acidity, Citric acid, Free sulfur dioxide, Total sulfur dioxide, Sulphates, Alcohol

- **Long Tail**: Residual sugar, Chlorides

### Rescale Variable

Skewed and long tail data can be transformed toward more normally distribution
by taking square root or log function. 
Take Sulphates as a example, we compare the original, square root and log of the feature.

```{r echo=FALSE,  Rescale_Variable}
# original 
p1<-
  ggplot(aes(x = sulphates), data = redwine)+
        geom_histogram(colour="black", fill="#ffb6c1", 
                      binwidth = 0.1)

# square root
redwine$sqrt_sulphates=sqrt(redwine$sulphates)
p2<-
    ggplot(aes(x = sqrt_sulphates), data = redwine)+
          xlab("square root of sulphates")+
          geom_histogram(colour="black", fill="#ffb6c1", 
                        binwidth = 0.05)

#log
redwine$log_sulphates=log(redwine$sulphates)
p3<-
   ggplot(aes(x = log_sulphates), data = redwine)+
          xlab("log of sulphates")+
          geom_histogram(colour="black", fill="#ffb6c1", 
                        binwidth = 0.1)

grid.arrange(p1,p2,p3,ncol=1)
```

### What is the structure of your dataset?

There are 1534 observations after slicing out the top 1% from the variables that had large outliers (Fixed acidity, residual sugar, total sulfur dioxide, and 
free sulfur dioxide)

### What is/are the main feature(s) of interest in your dataset?

Quality is the main feature. I want to determine what makes a wine taste good or bad.

### What other features in the dataset do you think will help support your analysis

### Did you create any new variables from existing variables in the dataset?

Yes, I created a rating variable which is a subset of quality based on three distinct categories: (bad: 4,5), (average: 5,6), (excellent: 7,8)

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

* The top 1% of values were stripped off of fixed acidity, residual sugar, total sulfur dioxide, and free sulfur dioxide.
* The x column was removed as it was simply an index value of unimportance.
* Sulphates appeared to be skwed and were log-transformed which revealed a 
normal distribution.

# Bivariate Plots Section

Plot matrix was used to have a glance at the data. We are interested the correlation between the wine quality and each chemical property.

```{r echo=FALSE, warning=FALSE, message=FALSE, Plot_Matrix}
# create a plot matrix
ggcorr(redwine, geom = "blank", label = TRUE, hjust = 0.75) +
  geom_point(size = 10, aes(color = coefficient > 0, alpha = abs(coefficient) > 0.5)) +
  scale_alpha_manual(values = c("TRUE" = 0.25, "FALSE" = 0)) +
  guides(color = FALSE, alpha = FALSE)

```

The top 4 factors that is correlated with the wine quality (with a correlation coeffcient greater than 0.2)

| Property         | r-value |
|------------------|---------|
| alcohol          | 0.49    |
| volatile.acidity | -0.39   |
| sulphates        | 0.256   |
| citric.acid      | 0.223   |

From the above table and plot matrix we see “fixed.acidity”, “volatile.acidity” 
and “pH” has some correlation with “citric.acid”. Interestingly, density has
some correlation with “fixed.acidity” and “alcohol”. Also, “quality” has some correlation with “alcohol”.

To see if the data makes sense chemically, I first plot pH and fixed acidity. 
The correlation coefficient is -0.68, meaning that pH tends to drop at fixed acidity increases, which makes sense.

```{r echo=FALSE, Bivariate_Plots_2}

# Create scatterplots for confirming relation with acidities and pH

# Fixed acidity vs pH

ggplot(aes(x = fixed.acidity, y = pH), data = redwine) +
  geom_point(color = "#F79420", alpha = 1/4) + 
  stat_smooth(method = 'lm')

cor(redwine$fixed.acidity, redwine$pH)
```

The correlation between citric acid and pH is slightly weaker, being -0.52.This adds up as citric acid is a subset of fixed acidity.

```{r echo=FALSE, Bivariate_Plots_3}

ggplot(aes(x = citric.acid, y = pH), data = redwine) +
  geom_point(color = "#F79420", alpha = 1/4) + 
  stat_smooth(method = 'lm')

cor(redwine$citric.acid, redwine$pH)

```


Volatile acidity (acetic acid) seems to increase when pH level increases. The correlation coefficient was 0.24 indicating some positive correlation.

```{r echo=FALSE, Bivariate_Plots_4}

ggplot(aes(x = volatile.acidity, y = pH), data = redwine) +
  geom_point(color = "#F79420", alpha = 1/4) + 
  stat_smooth(method = 'lm')

cor(redwine$volatile.acidity, redwine$pH)

```

I want to further explore alcohol, pH, volatile acidity, citric acid, and 
sulphates and see how they relate to the quality of the wine as they all had correlation coefficients greater than 0.2. Box plots are used and we use the
median as a better measure for the variance in our data. As predicted, the 
median also follows suit with the correlation coefficients. The boxplots provide 
an extremely interesting fact about alcohol -- alcohol content is significantly higher for excellent wines compared to bad or average wines. Sulphates and
citric acid also seem to be positively correlated to to quality, and volatile acidity appear to be negatively correlated.

```{r echo=FALSE, Bivariate_Plots_5}

# Create boxplots for variables of interest

ggplot(aes(x = rating, y = pH), data = redwine) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue') +
  stat_summary(fun.y = median, geom = 'point', shape = 4) +
  stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8,
               size = 4) +
  coord_cartesian(ylim = c(2.85, 3.75)) +
  labs(x = 'quality')
by(redwine$pH, redwine$rating, summary)

ggplot(aes(x = rating, y = alcohol), data = redwine) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue') +
  stat_summary(fun.y = median, geom = 'point', shape = 4) +
  stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8,
               size = 4) +
  labs(x = 'quality')
by(redwine$alcohol, redwine$rating, summary)

ggplot(aes(x = rating, y = volatile.acidity), data = redwine) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue') +
  stat_summary(fun.y = median, geom = 'point', shape = 4) +
  stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8,
               size = 4) +
  coord_cartesian(ylim = c(0.1, 1.2)) +
  labs(x = 'quality')
by(redwine$volatile.acidity, redwine$rating, summary)

ggplot(aes(x = rating, y = citric.acid), data = redwine) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue') +
  stat_summary(fun.y = median, geom = 'point', shape = 4) +
  stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8,
               size = 4) +
    coord_cartesian(ylim = c(0, 0.76)) +
  labs(x = 'quality')
by(redwine$citric.acid, redwine$rating, summary)

ggplot(aes(x = rating, y = sulphates), data = redwine) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue') +
  stat_summary(fun.y = median, geom = 'point', shape = 4) +
  stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8,
               size = 4) +
  coord_cartesian(ylim = c(0.2, 1.1)) +
  labs(x = 'quality')
by(redwine$sulphates, redwine$rating, summary)

```

However, none of the variables share much in common with alcohol - the highest
is pH, which had a correlation coefficient of 0.22. However, alcohol and quality have a 0.49 correlation coefficient, which may be leading me somewhere.

It appears that when citric acid is in higher amounts, sulphates are as well. 
The freshness from the citric acid and the antimicrobial effects of the 
sulphates are likely correlated. The correlation coefficient was 0.33 which indicates weak correlation, but still noteworthy.

```{r echo=FALSE, Bivariate_Plots_6}

# Plot citric acid vs log10 sulphates

ggplot(data = redwine, aes(x = log10(sulphates), y = citric.acid)) +
  geom_point(color = "#F79420", alpha = 1/2, size = 1, position = 'jitter') +
  stat_smooth(method = 'lm')

cor(log10(redwine$sulphates), redwine$citric.acid)
```

When graphing volatile acidity and citric acid, there is clearly a negative correlation between the two. It seems that fresher wines tend to avoid the use
of acetic acid. The correlation coefficient was -0.57, indicating that larger amounts of citric acid meant smaller amounts of volatile acidity. Since volatile acidity 
is essentially acetic acid, the wine makers would likely not put a large amount
of two acids in the wine, leading them to choose one or the other.

```{r echo=FALSE, Bivariate_Plots_7}

# Plot citric acid vs volatile acidity

ggplot(data = redwine, aes(x = citric.acid, y = volatile.acidity)) +
  geom_point(color = "#F79420", alpha = 1/2, size = 1, position = 'jitter') + 
  stat_smooth(method = 'lm')

cor(redwine$citric.acid, redwine$volatile.acidity)
```

There is no particularly striking relationship between alcohol and pH -- a weak positive correlation of 0.22.

```{r echo=FALSE, Bivariate_Plots_9}

# Plot alcohol vs pH

ggplot(data = redwine, aes(x = alcohol, y = pH)) +
  geom_point(color = "#F79420", alpha = 1/2, size = 1, position = 'jitter') +
  stat_smooth(method = 'lm')

cor(redwine$pH, redwine$alcohol)
```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

It appears that when citric acid is in higher amounts, sulphates are as well. 
The freshness from the citric acid and the antimicrobial effects of the
sulphates are likely correlated. Volatile acidity and citric acid are negatively correlated. It is likely that fresher wines avoid the bitter taste of 
acetic acid. Citric acid 
and pH were also negatively correlated -- a lower pH indicates a higher acidity.
pH and alcohol are very weakly correlated. Pure alcohol (100%) has a pH of 7.33,
so when it is diluted it will likely increase the pH level ever so slightly.

The boxplots reveal an interesting picture as well:

* The median for sulphates increased for each quality type. The biggest jump was from average to excellent, with a median of aproximately 0.74 for excellent and 0.61 for average.
* Citric acid had the highest concentration for excellent wines. The median
jumped evenly throughout the different quality categories. With medians of 0.075 for bad, 0.24 for average, and 0.395 for excellent.
* As volatile acidity increased, the median for the wine became worse, with 
medians of 0.68 for bad, 0.54 for average, and 0.37 for excellent. It's possible that past a certain threshold,the acetic acid became too bitter for the tasters.
* The median for alcohol content (10%) was the same the wine was bad or average. However, for the excellent wines, the alcohol content was 11.6%. This leads to a striking observation: a higher alcohol content may make a wine excellent from average, however there are other factors at play that make a wine taste bad.
* pH didn't change significantly much between the wines, with medians of 3.38
for bad, 3.31 for average, and 3.280 for excellent.

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

Volatile acidity and citric acid were negatively correlated, as were citric acid and pH. Fixed acidity and pH were negatively correlated, due to the lower pH 
more acidic effect.

### What was the strongest relationship you found?

From the variables analyzed, the strongest relationship was between Citric Acid 
and Volatile Acidity, which had a correlation coefficient of -0.563.`

# Multivariate Plots Section

### Main Chemical Property vs Wine Quality

With different colors, we can  add another dimension into the plot. There are 4 main features.Alcohol, volatile acidity are the top two factor that affect wine quality. 

```{r echo=FALSE, Multivariate_Plots}
  ggplot(aes(x = alcohol, 
           y = volatile.acidity  , color = factor(quality)), 
       data = redwine) +
      geom_point(alpha = 0.8, size = 1) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  scale_color_brewer(type='seq',
                   guide=guide_legend(title='Quality')) + theme_dark()

```

The figure looks over ploted, since the wine quality are discrete numbers. We
can use jitter plot to alleviate this problem

```{r echo=FALSE, Jitter_Plots}
  ggplot(redwine, aes(y=alcohol,x=quality,color = volatile.acidity) )+
        geom_jitter(alpha=1/2,size=2)+
        xlab("wine quality") + ylab("alcohol") +
        scale_colour_gradient2(low="red", high="blue",
                               mid="#9933CC",midpoint = 0.8)

```


We can see higher quality wine have higher alcohol and lower volatile acidity.

### Add Another Feature 

Now we add the third feature, the log scale of sulphates, and use different 
facet to show wine grade. 

```{r echo=FALSE, Third_Features, fig.width=7, fig.height=7}

ggplot(redwine, aes(x=alcohol,y=volatile.acidity)) +
      geom_point(aes(color =  log_sulphates),alpha=1/2,size=2)+
      xlab("alcohol") + ylab("log of sulphates") +
      scale_colour_gradient2(low="red", high="blue",
                            mid="#9933CC",midpoint = -0.25)+
  geom_smooth(method = "lm", se = FALSE,size=1)  +
      facet_grid(rating ~ .) 
```

We can see higher quality wine have higher alcohol (x-axis), lower volatile 
acidity (y-axis) and higher sulphates (hue).

### Main Chemical Properties  vs Wine Quality

Since we can visualized 3 dimensions, including wine quality, at a time. Two
graphs will be needed to visualize the 4 main chemical properties.

```{r echo=FALSE, Main_Chemical_Property}

ggplot(aes(x = volatile.acidity, y = alcohol), data = redwine) +
      xlab("volatile acidity") + ylab("alcohol") +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
      geom_point(aes(color = rating),size=1)+
  scale_color_brewer(type = 'seq')+ theme_dark()
```

The same trend of alcholand volatile acidity's effect on wine qaulity can be observed.

```{r echo=FALSE, Main_Chemical_Property2}

ggplot(aes(x = log_sulphates, y = citric.acid), data = redwine) +
      xlab("log sulphates") + ylab("citric acid") +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
      geom_point(aes(color =  rating),size=1) +
  scale_color_brewer(type = 'seq')+ theme_dark()
```

We can see higher quality wine have higher sulphates (x-axis), higher citric acidity (y-axis).

# Linear Multivariable Model

Linear multivariable model was created to predict the wine quality based on chemical properties. 

The features are selected incrementally in order of how strong the correlation between this feature and wine quality.  

```{r echo=FALSE, Linear_Model}

  # regression
  m1<-lm(quality ~ volatile.acidity,data=redwine)
  m2<-update(m1,~. + alcohol)
  m3<-update(m2,~. + sulphates)
  m4<-update(m3,~. + citric.acid)
  m5<-update(m4,~. + chlorides)
  m6<-update(m5,~. + total.sulfur.dioxide)
  m7<-update(m6,~. + density)
  mtable(m1,m2,m3,m4,m5,m6,m7)

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \investigation. Were there features that strengthened each other in terms of \looking at your feature(s) of interest?

Based on the multivariate analysis, five features stood out to me: alcohol, sulphates, citric acid, volatile acidity, and quality. Throughout my analysis, chlorides and residual sugar lead to dead ends. However, high volatile acidity 
and low sulphates were a strong indicator of the presence of bad wine. High 
alcohol content, low volatile acidity, higher citric acid, and lower sulphates 
all made for a good wine.

### Were there any interesting or surprising interactions between features?

Surprisingly, other chemical proprieties do not have strong correlation with wine quality, such as the residual sugar and PH .

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model

Yes, I created a linear model using seven variables: alcohol, citric acid, sulphates, volatile acidity, chlorides, total.sulfur.dioxide and density. The
model was less precise in predicting qualities of 3, 4, 7, and 8, where the
error was +/- 2. For qualities of 5 and 6, the majority of predictions were off
by 0.5 and 1 for each bound. The limitations of this model are obvious – I’m 
trying to use a linear model for data that obviously isn’t perfectly linear.

------

# Final Plots and Summary

### Plot One: Distribution of Wine Quality

```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}

ggplot(aes(x = quality), data = redwine)+
geom_bar(colour="black", fill="#ff9dab", binwidth = 1)+
                scale_x_continuous()+
                xlab("wine grade") + ylab("count") +
                ggtitle("Distribution of Wine Quality")

```

### Description One

The simplest but most informative, Plot shows the distribution of wine quality. 
Note that dataset is unbalanced and the distribution is mostly normal. 
It has many count for medium quality (grade 5, 6), but much fewer count on low (grade 3,4) and high (grade 7, 8) quality wine. 


### Plot Two: Alcohol & Sulphates vs. Quality & Volatile Acidity vs Quality

```{r echo=FALSE, Plot_Two}

# Plot 1: log Sulphates and citric acid
# Plot 2: Volatile acidity and alcohol

p1 <- ggplot(aes(x = log_sulphates, y = citric.acid), data = redwine) +
      xlab("log sulphates") + ylab("citric acid") +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
      geom_point(aes(color =  rating),size=1) + 
  scale_color_brewer(type = 'seq') + theme_dark()

p2 <- ggplot(aes(x = volatile.acidity, y = alcohol), data = redwine) +
      xlab("volatile acidity") + ylab("alcohol") +
  geom_smooth(method = "lm", se = FALSE,size=1)  +
      geom_point(aes(color = rating),size=1) + 
  scale_color_brewer(type = 'seq') + theme_dark()

grid.arrange(p1,p2,ncol=1)

```

### Description Two

The 4 features are also represented in the scatter plot. 2 features are plotted 
at a time with color indicate wine quality. Similar trend as the last figure can
be observed. In general, high quality wine tend to have higher alcohol and lower volatile acidity content. They also tend to have higher sulphate and higher
critic acid content.
As we can see, when volatile acidity is greater than 1, the probability of the
wine being excellent is zero. When volatile acidity is either 0 or 0.3, there is roughly a 40% probability that the wine is excellent. However, when volatile acidity is between 1 and 1.2 there is an 80% chance that the wine is bad. 
Moreover, any wine with a volatile acidity greater than 1.4 has a 100% chance of being bad. Therefore, volatile acidity is a good predictor for bad wines.

Observe that lower sulphates content typically leads to a bad wine.Average wines have higher concentrations of sulphates, however for citric content Excellent 
wines are increased with increase in concentration. 

### Plot Three: Boxplotting Main Features

```{r echo=FALSE, Plot_Three}

# Create scatter plot with two of the arguably most 
# important variable in determining bad wine.


p1 <- ggplot(aes(x = rating, y = alcohol), data = redwine) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue') +
  stat_summary(fun.y = median, geom = 'point', shape = 4) +
  stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8,
               size = 4) +
  labs(x = 'quality')
by(redwine$alcohol, redwine$rating, summary)

p2 <- ggplot(aes(x = rating, y = volatile.acidity), data = redwine) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue') +
  stat_summary(fun.y = median, geom = 'point', shape = 4) +
  stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8,
               size = 4) +
  coord_cartesian(ylim = c(0.1, 1.2)) +
  labs(x = 'quality')
by(redwine$volatile.acidity, redwine$rating, summary)

p3 <- ggplot(aes(x = rating, y = citric.acid), data = redwine) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue') +
  stat_summary(fun.y = median, geom = 'point', shape = 4) +
  stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8,
               size = 4) +
    coord_cartesian(ylim = c(0, 0.76)) +
  labs(x = 'quality')
by(redwine$citric.acid, redwine$rating, summary)

p4 <- ggplot(aes(x = rating, y = sulphates), data = redwine) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue') +
  stat_summary(fun.y = median, geom = 'point', shape = 4) +
  stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8,
               size = 4) +
  coord_cartesian(ylim = c(0.2, 1.1)) +
  labs(x = 'quality')
by(redwine$sulphates, redwine$rating, summary)

grid.arrange(p1,p2,p3,p4,ncol = 2)

```

### Description Three

This graph was interesting because it showed how excellent wines tended to have
a higher alcohol content all else equal. By this I mean certain precursors had
to exist for alcohol to be the predominant determininant for quality.
The 4 features that have the highest correlation coefficient are alcohol,
volatile acidity, sulphates, citric acid. The wine quality are grouped to 
low (3,4), medium (5.6) and high (7,8).High quality wine have high alcohol level however, there is no significant different between medium and low quality wine. Volatile acidity decrease as wine quality increases. Sulphate and critic acid increase as wine quality increase.

------

# Reflection

This analysis was conducted conducted with the view of trying to uncover hidden insights by move a step at a time and proceeding further or retracting backwards based on the outcome. It was at times unbelievable at times when the hypothesis
was incorrect, but it did make sense. The most important thing that influenced 
the direction on the analysis was some sort of patterns that unravelled.

The biggest struggle in this process was working though the number of iterations needed to get the results out correctly, which in itself is a very tedious
process. I felt like giving up at times, but instead I decided to work through 
it one step at a time.

In the future analysis, it would make sense to carry out analysis based on the 
free radicals.

The take aways from this analysis are that wines with high quality tend to have higher alcohol content and low residual sugar. Another interesting finding was 
that citric acidity decreases with pH levels. So, wines with lower acidty have higher citric acid content.

In conclusion, if you are looking for a good bottle of wine. It will most like 
have very little sweetness to it, and good amount of alchohol.

### References
https://www.udacity.com

https://en.wikipedia.org

https://www.stackoverflow.com